### 2025-05-19 src/for_reviewers.tex

\documentclass{article}
\usepackage{pdfpages}
\usepackage{graphicx}
\begin{document}

\includepdf[pages=-]{rebuttal.pdf}
\section*{Changes made}
\includepdf[pages=-]{diff.pdf}
\section*{New version}
\includepdf[pages=-]{main.pdf}
\end{document}