### 2025-05-20 src/for_reviewers.tex

\documentclass{article}
\usepackage{pdfpages}
\usepackage{graphicx}
% \begin{document}
\title{%
  SHORT: Longitudinal Study of Software Environments Produced by Dockerfiles from Research Artifacts: Initial Design
	\\
	{\Large%
	Answer to reviewers
	}
}

\author{Anonymous Authors}
\date{}

\begin{document}

\maketitle

\begin{abstract}
This document is composed of three parts:

\begin{enumerate}
\item the detailed answers to the reviewers comments
\item a ``diff'' between the initial submission and the new submission (additions in underlined blue, deletions in crossed-out red)
\item the new submission of the article
\end{enumerate}
\end{abstract}

\tableofcontents

\section{Answers to reviewers}
\includepdf[pages=-]{rebuttal.pdf}
\section{Changes made}
\includepdf[pages=-]{diff.pdf}
\section{New version}
\includepdf[pages=-]{main.pdf}
\end{document}